\documentclass[11pt, a4paper]{article}

\usepackage{amsmath, amssymb, enumerate}

\title{Sugggestions for HW6}

\begin{document}

Please use the Poisson in the homework and the geometric in the exercises sheet. I think they need to see more Poissons.

\section{Poisson EM}

You observe the data set $ x = \{3, 6, 8, 3, 5, 4, 7, 2, 2, 8\} $. You decide to model with a mixture model with 5 components $ c_{1}, c_{2}, c_{3}, c_{4}, c_{5} $ which
are linked to Poisson distributions with parameters $ \lambda_{c_{1}} = 0.5, \lambda_{c_{2}} = 1, \lambda_{c_{3}} = 1.5, \lambda_{c_{4}} = 2, \lambda_{c_{5}} = 3 $. 
Initially you assume that all the mixture components are equally likely. You
also assume independence between the mixture components. Thus, your model is
\begin{equation*}
P(X_{1}^{10}=x_{1}^{10},Y_{1}^{10}=y_{1}^{10}|\Lambda=\lambda) = \prod_{i=1}^{10}P(Y_{i}=y_{i})P(X_{i}=x_{i}|Y_{i}=y_{i},\Lambda=\lambda)
\end{equation*}
where the $ x_{i} $ are observed and the $ y_{i} $ are latent. Each $ Y_{i} $ can take on values in $ \{ c_{1}, c_{2}, c_{3}, c_{4}, c_{5} \} $.

\begin{enumerate}[a)]
\item Compute the marginal log-likelihood of $ x $ under the initial parameter settings.
\item Perform the E-step on the data. Compute the posterior probabilities for the mixture components per data point.
\item Report the expected sufficient statistics $ \mathbb{E}[c=c_{i}|\lambda] $ for each $ 1 \leq i \leq 5 $. 
\item Report the expected sufficient statistics $ \mathbb{E}[x=v|c_{i},\lambda] $ for $ v \in \{2,3,4,5,6,7,8\} $. Use a table for this exercise!
\item Perform the M-step. Report the new parameters of the distributions $ P_{Y} $ and $ P_{X|Y_{i}} $ for  $ 1 \leq i \leq 5 $.
\item Compute the marginal log-likelihood of of $ x $ under the new parameter settings. \textbf{Hint:} If you have done everything correctly, this value should
be higher then what you got under the initial parameter settings.
\end{enumerate}

\end{document}