\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{calc}
\usepackage{fullpage}
\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue, breaklinks=true}

%\newcommand{\prop}[1]{\lbrack \! \lbrack #1 \rbrack \! \rbrack}

\newcommand{\philip}[1]{ \textcolor{red}{\textbf{Philip:} #1}}
\newcommand{\chris}[1]{ \textcolor{blue}{\textbf{Chris:} #1}}


\title{Theory Assignment 5 -- Basic Probability, Computing and Statistics\\[2mm]
\large{Fall 2015, Master of Logic, University of Amsterdam}}

\author{}
\date{Submission deadline: Monday, October 12th, 2015, 9 a.m.}



\begin{document}
\maketitle

\paragraph{Cooperation}
Cooperation among students for both theory and programming exercises
is strongly encouraged.  However, after this discussion phase, every student writes down and submits his/her own individual solution.

\paragraph{Guidelines}
You may pick {\bf N exercises from exercise type I}, as well as {\bf M from exercise type II} for submission, i.e. you need to submit {\bf a total of N+M exercises} to be able to get all points. Numbered exercises with an exclamation mark are supposed to be a bit harder and you may challenge yourself by trying to solve them.

In the directory of your private url there is folder called `theory\_submissions'. Please upload your submission there. Your submission should be a PDF-document (use a scanner for handwritten documents!) entitled \textit{AssignmentX\_yourStudentNumber.pdf}, where \textit{X} is the number of the assignment and \textit{yourStudentNumber} is your student number. If your submission does not comply with this format, we will deduct 1 point. For each day that your submission is late, we deduct 2 points. N.B.: If multiple files are submitted for a single assignment before the deadline, the latest version will be graded.

If you have any question about the homework or if you need help, do not hesitate to contact \href{mailto:T.S.Brochhagen@uva.nl}{Thomas}.

\paragraph{Exercises}

\paragraph{Type I [N exercises: Q point per exercise]}
\begin{enumerate}
	\item Consider the outcome of (i) a flip of a fair coin, (ii) a toss of a fair four-sided die, and (iii) a toss of a fair six-sided die. Let a RV $X$ encode (i), (ii) and (iii) and compute $H(X)$ in each case.
	\item A biased coin comes up heads with a probability of $\frac{2}{3}$. Compute the entropy of the outcome of six coin flips.
	\item Let $X$ encode the sum of the roll of two fair dice. Compute $H(X)$.
	\item Assume that there is an equal probability of $50\%$ for a person in a population to be male or female. Suppose further that $20\%$ of the males and $6\%$ of the females are tall (height greater than some fixed threshold). Calculate the information gained by (i) learning that a male person is tall, (ii) a female person is tall, (iii) a tall person is female.
\end{enumerate}

\paragraph{Type II [M exercises: R points per exercise]}
\begin{enumerate}
	\item Show that relative entropy does not satisfy (i) symmetry nor (ii) triangle inequality. That is, it needs not hold that $D(X\parallel Y) \neq  D(Y \parallel X)$ nor $D(X\parallel Y) + D(Y\parallel Z) \geq D(X \parallel Z)$.
	\item Show that $D(X\parallel Y) \geq 0$.
	\item Prove that the information gained from observing $n$ independent RVs is the sum of the information gained from their individual observation, i.e. that  $H(X_1, ..., X_n)$ is additive if $P(\bigcap_{i = 1}^{n} X_i=x) = \Pi_{i = 1}^n P(X_i=x)$.
\end{enumerate}
\end{document}
