\documentclass[11pt, leqno, a4paper]{article}
\usepackage{hyperref, amsmath, amssymb}
\hypersetup{colorlinks=true, urlcolor=blue, breaklinks=true}

\title{Programming Assignment 5 -- Basic Probability, Computing and Statistics 2016 \\[2mm]
\large{Fall 2016, Master of Logic, University of Amsterdam}}
\author{Instructors: Philip Schulz, Christian Schaffner and Bas Cornellisen}
\date{Submission deadline: Wednesday, December 7th, 8 p.m.}

\begin{document}
\maketitle

\paragraph{Note:} if the assignment is unclear to you or if you get
stuck, do not hesitate to ask in the
\href{https://www.moodle.ch/lms/mod/forum/view.php?id=1721}{forum}.

\section{Assignment}

This week we are finally going to implement the expectation maximisation algorithm (EM). You can look up explanations of the algorithm 
\href{https://github.com/BasicProbability/LectureNotes/blob/master/chapter6/chapter6.pdf}{here} and 
\href{}{here}. Our data (download \href{https://github.com/BasicProbability/BasicProbability.github.io/raw/master/Homework/Programming/2016-17/Assignment5/geometric_data.txt}{here}) have been generated from a mixture model. The mixture components are geometric distributions. Your tasks is to find
the mixture weights and parameters of the mixture components. The data were generated from 5 mixture components, so this is the number of components
that you should use as well.

Recall that EM is sensitive to the starting point. During peer review, we will
provide you with a starting point whose resulting parameter estimates we have already computed. You will be assessed according to how well your
algorithm does on this particular starting point. This implies that you will have to provide a place in your code (ideally at the beginning) where
the reviewers can plug in that fixed starting point.

\section{Implementation}

We do not provide skeleton code this time. You will have to write all functions (and classes if you want to work with those) from scratch. You are encouraged
to discuss your implementation in the forum. While you should not post any code, asking questions like \textit{Do you think that this is the right data structure
to use?} are is perfectly fine. 

Notice that we are again in a setting were we have to work with logprobs. In order to compute the log-likelihood (see next section), we have to add those logprobs.
This is where the logarithm functions that we implemented in week 3 come in handy. Please use them whenever needed. We provide a correct implementation of
them \href{https://github.com/BasicProbability/BasicProbability.github.io/raw/master/Homework/Programming/2016-17/Assignment5/logarithms_solution.py}{here}.

\section{Debugging}

Recall that EM is guaranteed to always increase the log-likelihood of the data. Therefore, you should compute and print the log-likeihood after each iteration.
If it goes up after each iteration or stays unchanged, your implementations is probably correct (no guarantees, though). Please make sure to be very strict about
this. Even if an EM implementation is wrong, the log-likelihood often still increases during the initial iterations. Pay close attention to later iterations. If they
are volatile, you have a bug! EM usually does not need all too many iterations before it converges. Always run 20 iterations, which should be more than enough for 
this data set.

\section{Grading}

You will be graded on your performance from a specific starting point, i.e. a specific setting of initial mixture weights and parameters of mixture components.
Make sure that these can be provided to your code. This can be done either by defining list variables at the beginning of your python file (one list of the
mixture weights, one list for the component parameters) or, more elegantly, by asking the user to provide a text file in a specific format from which your program
reads the initial parameter settings.

\begin{itemize}
\item[1 point] All classes and functions/methods have docstrings. Award 0 points here if there are one or more classes/functions/methods that do not have a 
docstring.
\item[1 point] A starting point for EM can be provided to program.
\item[1 point] If no starting point is provided, the parameters are initialised randomly.
\item[1 point] The program prints the log-likelihood after each iteration. After iteration $ t $ it is fine to print the likelihood achieved with the parameter
settings from  iteration $ t-1 $. This simplifies the computation (figure out yourselves how).
\item[3 points] The log-likelihood increases monotonically when the algorithm is run.
\item[3 points] Your parameter estimate after 20 iterations for the fixed initial parameter settings is equal to the parameter estimate that we will provide during the 
peer review period.
\end{itemize}


\end{document}